### 2022-02-11 get_data.Rmd
testing examples

---
title: "Get Data"
author: "Erin M. Buchanan"
date: "Last Knitted: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(dplyr)
library(purrr)
```

## Crossref

Things we need to do to update:

  - We need to put together a list of dois to pull from crossref.

Things you need to do once:

  - Install crossref from github because it has an error that hasn't been pushed to cran.
  - Add your creds to the environment file so it knows who you are.

```{r}
library(rcrossref)
# remotes::install_github("ropensci/rcrossref") as of 12/2021 because of error in crul
# be nice to crossref
# file.edit("~/.Renviron")
# add crossref_email= "email@gmail.com"

# example, here we would loop over the dois 
# psa001 <- cr_cn("10.1038/s41562-020-01007-2") bug in crossref geez

osc <- cr_cn("10.1126/science.aac4716")

manylabs1 <- cr_cn("10.1027/1864-9335/a000178")

bibs <- bib2df("papers.bib")
```

## ORCID

Things you need to do one time:

- Use `orcid_auth()` to set you up
- Copy the bearer number and put it in your environment file

```{r}
library(rorcid)
# orcid_auth() run the first time to set up
# copy the bearer number 
# file.edit("~/.Renviron")
# ORCID_TOKEN="numberhere"

orcid_search(given_name = 'erin',
             family_name = 'buchanan')

erin_example <- orcid_person("0000-0002-9689-4189")

erin_df <- rorcid::works("0000-0002-9689-4189") %>%
  as_tibble() %>%
  janitor::clean_names() %>%
  dplyr::mutate(created_date_value = anytime::anydate(created_date_value/1000))

```

May need to manually edit some stuff
Figure out how to get information

Ideas from meeting:
- Career length: from first pub, from first degree
- Over time: are things becoming more varied
- Maybe we can use scholar if they don't have orcid
- Which journals are publishing this information
- Types of big team research (social, cognitive)
- Employment levels
- Education levels
- # Of works: separate to total, other/preprint, journal/book/traditional
- Limitations of this system