### 2022-02-25 get_data.Rmd
updated id files

---
title: "Get Data"
author: "Erin M. Buchanan"
date: "Last Knitted: `r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(dplyr)
library(purrr)
library(bib2df)
library(tidyverse)
library(httr)

library(rorcid)
# orcid_auth() run the first time to set up
# copy the bearer number 
# file.edit("~/.Renviron")
# ORCID_TOKEN="numberhere"

library(scholar)
```

## Import the BibTeX File

In this section, we are importing the bibtex file that includes all the papers we are going to use.

Notes:

  - We used the preprint manuscript for some papers because they had the entire author list, we will update their references manually for the final dataset. 
  - We will have to manually add the names for several papers that used consortium authors. 
  - The pre-reg of this document only examines one paper to ensure code can be run and develop workflow.

```{r}
bibs <- bib2df("papers.bib")
```

## ORCID

In this section, we will figure out the ORCIDs for each author if possible.

  - Data is stored in `bibs$AUTHOR` as a list. 
  - Extract that information and loop over the author list. 
  - Because authors will repeat, we need to put this in long format with one row per author and all other information saved. 
  - Next, split author name, so we can look it up with orcid.

```{r}
# get data into long format
long_bib <- bibs %>% unnest_longer(col = AUTHOR)

# remove this nonsense
long_bib$AUTHOR <- gsub("[{}]", "", long_bib$AUTHOR)

# now split the names into last, first middle 
# first split on the comma only 
long_bib <- long_bib %>% separate(col = AUTHOR, sep = ",", 
                                  into = c("last", "first"), 
                                  extra = "merge")

# remove the first space
long_bib$first <- gsub("^ ", "", long_bib$first)

# then split on space
long_bib <- long_bib %>% separate(col = first, sep = " ", 
                                  into = c("first", "middle"),
                                  extra = "merge")

# drop editor column so it is a non-nested DF
long_bib <- long_bib %>% select(-EDITOR)
write.csv(long_bib, "unedited_bib.csv", row.names = F)
```

To make the finalized data:

  - Loop over rows and find candidates for ORCIDs.
  - Export the candidate data.
  - Export the bibtex data (done above).
  - A bit of fun manual merging.

```{r eval = F}
# there are going to be a lot of duplicates
# really only need to check them once if exact match
# not perfect exclusion because of symbol differences and middles but ok 
get_id <- long_bib %>% select(last, first, middle, TITLE)
get_id <- get_id %>% filter(!duplicated(get_id[ , c("last", "first")]))

# clean up titles
get_id$TITLE <- gsub("[[:punct:]]", "", get_id$TITLE)
get_id$TITLE <- tolower(get_id$TITLE)

get_id$ORCID <- NA

for (i in 1:nrow(get_id)){
#for (i in 1:10){ #temporary do a small number

  # find possible candidates 
  possibles <- orcid_search(family_name = get_id$last[i], 
                            given_name = get_id$first[i])

  cat(i)
  cat("\n")

  # if nrow = 0 then move on

  # if nrow == 1 then figure out if there's a match
  if (nrow(possibles) == 1){
    temp_works <- orcid_works(possibles$orcid) 
    temp_DF <- do.call(bind_rows, temp_works)

    if (nrow(temp_DF) > 0){
    temp_DF$title.title.value <- gsub("[[:punct:]]", "", temp_DF$title.title.value) 
    temp_DF$title.title.value <- tolower(temp_DF$title.title.value) 
    temp_DF <- temp_DF %>% 
      filter(temp_DF$title.title.value == get_id$TITLE[i])

    id <- unique(na.omit(temp_DF$`source.assertion-origin-orcid.path`))

    if(length(id) == 1) { get_id$ORCID[i] <- id }
    if(length(id) > 1) { get_id$ORCID[i] <- list(id) }

    }
  }

  # if nrow > 1 figure it out which matches
  if (nrow(possibles) > 1){
    temp_works <- orcid_works(possibles$orcid) 
    temp_DF <- do.call(bind_rows, temp_works)

    if (nrow(temp_DF) > 0){
    temp_DF$works$title.title.value <- gsub("[[:punct:]]", "", temp_DF$works$title.title.value) 
    temp_DF$works$title.title.value <- tolower(temp_DF$works$title.title.value) 
    temp_DF <- temp_DF %>% 
      filter(temp_DF$works$title.title.value == get_id$TITLE[i])

    id <- unique(na.omit(temp_DF$works$`source.assertion-origin-orcid.path`))

    if(length(id) == 1) { get_id$ORCID[i] <- id }
    if(length(id) > 1) { get_id$ORCID[i] <- list(id) }

    }
  }

  # be nice to api take a break 
  Sys.sleep(runif(1, 3, 5))

}

write.csv(get_id, "possible_orc_ids.csv", row.names = F)
```

## Google Scholar

- First, fix the scholar package issue that it doesn't correctly pull ids that have punctuation in them.

```{r eval = F}
get_scholar <- function (last_name = "", first_name = "", affiliation = NA) 
{
    if (!any(nzchar(c(first_name, last_name)))) 
        stop("At least one of first and last name must be specified!")
    site <- getOption("scholar_site")
    url <- paste0(site, "/citations?view_op=search_authors&mauthors=", 
        first_name, "+", last_name, "&hl=en&oi=ao")
    page <- get_scholar_resp(url)
    if (is.null(page)) 
        return(NA)
    aa <- content(page, as = "text")
    ids <- stringr::str_extract_all(string = aa, pattern = ';user=.{12}')

    if (length(unlist(ids)) == 0) {
        message("No Scholar ID found.")
        return(NA)
    }
    ids <- ids %>% unlist %>% gsub(";user=", "", 
        .) %>% unique

    return(ids)
  }
```

- Then determine if you can find scholars matching that information.

```{r eval = F}
# thanks SO for this id for the bad coding on scholar 
p_get_scholar_id <- possibly(get_scholar, otherwise = NA_character_)
scholars <- character(nrow(get_id))

for(i in 1:length(scholars)) {

  id <- NA
  possibles <- p_get_scholar_id(last_name = get_id$last[i], 
                                first_name = get_id$first[i])

  for (q in 1:length(possibles)){

    articles <- get_publications(possibles[q])
    articles$title <- gsub("[[:punct:]]", "", articles$title) 
    articles$title <- tolower(articles$title) 
    temp_DF <- articles %>% 
      filter(title == get_id$TITLE[i])

    if (nrow(temp_DF) > 0){

      if (!is.na(id)){ 
        id <- c(id, possibles[q]) 
        } else { id <- possibles[q] }
    }

  }
  cat(i)
  cat("\n")

  if (length(id) == 1) { scholars[i] <- id }
  if (length(id) > 1) { scholars[i] <- list(id) }

  # be nice to api take a break 
  Sys.sleep(runif(1, 3, 5))
}

get_id$scholar <- scholars

write.csv(get_id, "possible_both_ids.csv", row.names = F)
```

## Get Person Statistics

```{r}
erin_df <- rorcid::works("0000-0002-9689-4189") %>%
  as_tibble() %>%
  janitor::clean_names() %>%
  dplyr::mutate(created_date_value = anytime::anydate(created_date_value/1000))

table(erin_df$type)

erin_df$title_title_value <- tolower(erin_df$title_title_value)
erin_df$title_title_value <- gsub("[[:punct:]]", "", erin_df$title_title_value)
nrow(erin_df)

# remove real duplicates
temp <- erin_df[!duplicated(erin_df[ , c("type", "title_title_value") ]), ]
nrow(temp)

# combine types
erin_df$our_type <- erin_df$type
erin_df$our_type <- gsub("other|data-set|preprint", "open_science_pubs", erin_df$our_type)

table(erin_df$our_type)

erin_ed <- orcid_educations("0000-0002-9689-4189") %>% 
  as_tibble() %>%
  janitor::clean_names()

orcid_employments("0000-0002-9689-4189")

```

May need to manually edit some stuff
Figure out how to get information

Person statistics: 
- Career length: from first pub, from first degree
  `publication_date_year_value` anything 
- Employment levels
- Education levels
- # Of works: separate to total, other/preprint/data-set, journal/book/traditional
- Limitations of this system

...average citation rates/h index/i10 index

Journal statistics: 
- Which journals are publishing this information
  - make sure preprints we are using also have real journal merged in 
- Types of big team research (social, cognitive)
  - keywords 
- Over time: are things becoming more varied
  - are the publications more varied over time