### 2023-08-09 data/combine_outputs.Rmd

---
title: "Combine Files"
author: "Erin M. Buchanan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
library(rio)
library(dplyr)
```

## Combinations

```{r}
## geo political region 
# row	affiliation_tag_country	count	all_count
# 0	None	128448	1201063
# geo_country.xlsx

geo1 <- import("data/scopus_combine/geo_country_1.csv")
geo2 <- import("data/scopus_combine/geo_country_2.csv")

geo_together <- geo1 %>% 
  full_join(
    geo2, by = c("affiliation_tag_country" = "country")
  )
colnames(geo_together) <- c("affiliation_tag_country", "count", "all_count")

geo_together$count[is.na(geo_together$count)] <- 0

export(geo_together, "data/scopus_outputs/geo_country.xlsx", rowNames = F)
```