### 2023-08-09 manuscript_scopus.Rmd

---
title             : "Who does big team science?"
shorttitle        : "Big Team Science"
author: 
  - name          : "Erin M. Buchanan"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "326 Market St., Harrisburg, PA 17101"
    email         : "ebuchanan@harrisburgu.edu"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Data curation
      - Formal Analysis
      - Methodology
      - Project administration
      - Visualization
      - Writing – original draft
      - Writing – review & editing
  - name          : "Savannah C. Lewis"
    affiliation   : "2"
    role:
      - Conceptualization
      - Data curation
      - Methodology
      - Project administration
      - Writing – original draft
      - Writing – review & editing
affiliation:
  - id            : "1"
    institution   : "Harrisburg University of Science and Technology"
  - id            : "2"
    institution   : "University of Alabama"
authornote: |
  Erin M. Buchanan is a Professor of Cognitive Analytics at Harrisburg University of Science and Technology. Savannah C. Lewis is a graduate student at the University of Alabama.

  Thank you to Dwayne Lieck for providing an extensive list of large scale projects for this manuscript. 
abstract: |
  This paper examined the nature of publications in Big Team Science (BTS) - large-scale collaborations between multiple researchers at multiple institutions. As interest in BTS increases, it is useful to explore who is currently involved in BTS projects to determine diversity in both research subject and researcher representation. The types of publication outlets, number of publications, and subject areas of publication are presented to summarize the publications in BTS. Information about authors included in BTS will be presented including career length, numbers of publications/impact variables, education, and affiliation. Last, we will explore the representation of geopolitical regions by examining affiliation location to explore the impact of BTS on the de-WEIRD movement to diversify researcher representation. REWRITE THIS

keywords          : "big team, science, authorship, credit"
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
classoption       : "man"
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: inline
bibliography: references.bib
# header-includes:
#   - |
#     \makeatletter
#     \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
#       {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
#       {-1em}%
#       {\normalfont\normalsize\bfseries\typesectitle}}
#       {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
#       {-\z@\relax}%
#       {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
# csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
# documentclass     : "apa7"
output            : papaja::apa6_pdf
appendix          :
  - "appendix.Rmd"
---

```{r setup, include = FALSE}
knitr::opts_knit$set(echo = FALSE, warning = FALSE, message = FALSE, include = FALSE, error = FALSE)
library(papaja)
library(knitr)
library(ggplot2)
library(treemapify)
library(countrycode)
library(rio)
library(maps)
library(dplyr)
library(tidyr)
library(MOTE)
library(MBESS)
library(patchwork)
library(ggplot2)
library(ggwordcloud)
library(figpatch)
library(RColorBrewer)
library(latex2exp)

strip_labels <- expression(
  "M Coefficient " = b[M],
  "SD Coefficient" = b[SD],
  "Effect Size" = R^2,
  "Coefficient" = b
)

label_expressions <- function(values) {
  stopifnot(is.expression(values))
  as_labeller(function(x) {
    if (is.null(names(values))) {
      x <- seq_along(x)
      if (length(x)!=length(values)) warning(paste0("Number of labels(", 
        length(values), ") does not match number of values (", length(x), ")"))
    }
    as.list(values[x])
  }, default = identity)
}

```

```{r data, include = FALSE}
# things to import
var_list <- c("author_time", "career_time", "geo_country", "inst_time", 
              "keywords", "large_first_gp", "large_other_gp", "medium_first_gp", 
              "medium_other_gp", "normal_pub_type", "pub_time_count", 
              "pub_time_subject", "pub_types", "small_first_gp", "small_other_gp",
              "snip", "ASJC_codes", "snip_all")

var_list_import <- paste0("data/scopus_outputs/", var_list, ".xlsx")

DF <- lapply(var_list_import, import)
names(DF) <- var_list
```

According to the Oxford English dictionary, collaboration is two or more
people working together to achieve a certain goal [@oed2016].
Collaboration in scientific endeavors involves multiple researchers at
(potentially) multiple institutions to communicate and work together to
advance knowledge in their chosen field. Collaboration can manifest
uniquely in each project dependent on the skill sets, hypotheses, and
perspectives of collaborators. While collaboration is not new in
science, the current interest of "big team science" is increasing
[@forscher2020; @stewart2017; @coles2022]. Big team science projects
and/or organizations utilize and run on large-scale collaboration to
ensure that diverse populations and ideas are brought into research
projects, which in turn allows for more reliability and generalizability
in the results and method of the study. For this study, Big Team Science
(BTS) will be defined as a collaboration of ten or more authors from at
least ten different institutions.

BTS appears to be increasing as a result of two sources: 1) increasing
globalization and technology that allows for real-time interdisciplinary
research, and 2) increasing interest in reproducibility, replication,
and generalizability [@maxwell2015; @nelson2018; @zwaan2018].
Technological advances have provided easier ways to collaborate with
people who are from other universities and countries through document
sharing platforms (e.g., Google, GitHub, and the Open Science
Framework), video chatting platforms (e.g., Zoom, Microsoft Teams), and
messaging and project management platforms (e.g., Slack, Trello,
when2meet, etc.). The credibility movement seems to suggest that by
having both collaborations that span across the globe and subfields of
research areas, age groups, and education levels should help to drive science in the path of better materials, reliability,
generalizability and more robust sample sizes (when necessary) in a study
[@auspurg2021has; @nosek2014method; @lebel2018].

The credibility movement was originally defined by a focus on large
scale replications using in collaborative environments [@vazire2022].
Generally, the movement appears to have been driven by early career researchers
(i.e., those who are within five years of their first appointment)
[@maizey2019]; however, there are no large meta-scientific
investigations on this specific topic to date. Potentially, the lack of
investigation is tied to the newness of the large-scale research in many
fields, as it is only in recent years that publications like the Open
Science Collaboration [@opensciencecollaboration2015], Many Labs
Collaborations [for example, @klein2018; @ebersole2020; @ebersole2016;
@klein2022; @mathur2020; @buttrick2020; @skorb2020] or the first papers
from the Psychological Science Accelerator [@moshontz2018; @jones2021;
@dorison2022; @wang2021; @legate2022; @bago2022]. Generally, the
researcher incentive for replication was low: journals often prioritize
"novel" or new results which led to rejection of replication manuscripts
and publication bias [@hubbard1997; @franco2014; @nosek2012], the
"failure" to replicate was often placed on the replication team as "bad
science" rather than a careful consideration of publication biases and
(potential) questionable research practices [@ioannidis2015;
@maxwell2015; @klein2022], and why should someone want to spend time and
resources on an answer we already "know" [@isager2021; @isager2021a]?

However, the success and interest in the large-scale reproducibility
projects [@opensciencecollaboration2015; @10.7554/eLife.71601], paired
with the meta-scientific publications focusing on researcher practices
and incentive structures [@silberzahn2018many; @john2012] led to a
change in journal guidelines and incentives for researchers interested
in participating in large-scale replication studies [@grahe2014;
@kidwell2016; @nosek2015; @mayo-wilson2021]. For example, the support
for Registered Reports, papers accepted before the data has been
collected [@nosek2014; @stewart2020], and entire sub-sections of
journals devoted to only replication studies (e.g., *Nature, Royal Society Open Science, Advances in Methods and Practices in Psychological Science*) has allowed researchers to invest in projects that they know
should be published when the project is complete. Further, the
implementation of the Transparency and Openness Guidelines [@nosek2015]
and the Contributor Role Taxonomy (CRediT) system [@allen2019] have
pushed journals and researchers to promote more open, inclusive
publication practices.

The credibility movement has been mirrored by the calls for
diversification or de-WEIRDing (e.g., Western, Educated, Industrialized,
Rich, and Democratic) scientific research [@henrich2010; @newson2021;
@rad2018] by improving representation in research samples. Like the
large-scale studies in Physics [@aphilos2021; @castelnovo2018] and
Biology [@collins2003], the social sciences struggle to represent the
breadth of humanity across both researcher and population
characteristics. Now, grassroots organizations, such as the
Psychological Science Accelerator [@moshontz2018], ManyBabies
(<https://manybabies.github.io/>), NutNet (<https://nutnet.org/>), and
DRAGNet (<https://dragnetglobal.weebly.com/>) can begin to tackle these
issues by recruiting research labs from all over the globe to provide
diversity in geographic, linguistic, and researcher representation.
Publications have examined the global understanding of morality, face
processing, COVID-19 information signaling, and more [@dorison2022;
@wang2021; @legate2022; @bago2022; @jones2021; @vanbavel2022]. While
these organizations and one-time groups for BTS studies have provided an
incredible wealth of data for the scientific community, we do not yet
know exactly *who* is involved with, and benefits from, the BTS and
credibility movement. Publications on BTS generally explore challenges,
lessons learned, and the need for BTS [@forscher2020; @coles2022].

Therefore, the goal of this manuscript is to examine the *people*
involved in BTS projects. We specifically examined the themes of inclusivity, research careers, and research globalization. We see an
increasing interest and number of publications in BTS but we do not yet know if
this uptick in large-scale projects has diversified the *people*
involved in BTS. While a few publications have noted that BTS appears to
be early career researchers [@maizey2019], no one has systematically
investigated this perception. Further, it is unclear if the focus of
de-WEIRDing science has only focused on the representation of the
research participants or if it has also improved the representation of
researchers outside of North America and Europe. Last, who runs these
BTS projects? Do we see an increase in diversity for the authors who
generally receive the most credit for these projects (i.e., first
several author(s) and last author)? As hiring and promoting practices
often place a heavy weight on publications and especially "influential"
publications, it becomes necessary to critically examine the
representation present in authorship in BTS projects.

# Research Questions

-   Research Question 1: What publication sources publish big team
    science papers?
-   Research Question 2: What are the types of articles that are being
    published in big team science?
-   Research Question 3: Who is involved in big team science?

This manuscript was preregistered with the same conceptual ideas using Google Scholar and ORC-ID databases (https://osf.io/f2dtr) but then was updated with access to the Scopus database for a broader picture of BTS projects (https://osf.io/fheun). All materials and code can be found on our OSF page: https://osf.io/cgx6u/ or corresponding GitHub archive: https://github.com/doomlab/big_team_who.

# Method

## Publications

We have defined BTS publications as publications with at least ten
authors at ten different institutions that were published in
peer-reviewed journals or had posted a full paper pre-print. We used
data from 1970 and forward in the Scopus database, as it
is noted online that this time period includes cited references for
calculation of several of our variables described below. We will analyze
our results based on four subject areas present in the Scopus database: Physical Sciences,
Health Sciences, Social Sciences, and Life Sciences. We filtered the
database to include articles, articles in press, business articles,
conference papers, data papers, preprints, and surveys using Elsevier's
classification system. This project was supported by access to the Scopus database through the International Center for the Study of Research.

## Data Curation

### RQ1: Publisher Information.

We extracted the following information for publication sources: the name of the publication (source title), subject area (both the large four subject areas and the smaller four digit all science journal classification [ASJC] code), and the journal
impact using the Source Normalized Impact per Paper (SNIP).

### RQ2: Publication Information.

For each publication of the identified BTS publications, we examined
the full four digit ASJC subject areas codes for each of the larger four
subject areas and the keywords present for these publications.

### RQ3: Author Information.

The author list was extracted from each publication. Next, we used the author and affiliation arrays to curate a list of all publications and author information included in BTS papers to calculate the variables described below.

***Career Length***. Career length for each author was defined as
the year of the first publication minus the current year listed for each author.

***Institution and Geopolitical Region***. We used the affiliation
ids and country to gather information about the places of education
and/or employment for authors. Geopolitical region was created by binning these codes into United Nation Regions.

***Education***. We collected degree information from the author table.

***Types of Publications***. We took information from the
publication type variable for each author's publications to present
information about the types of papers BTS authors publish.

***Publication Metrics***. For each author, we calculated the total
number of publications, and the h-index. The h-index represents the
highest *h* number of publications that have at least *h* citations.

# Results

We used the 95% confidence interval to make decisions on predictor or effect size differences from zero. The confidence interval that does not include zero would be considered different from zero (to four decimal places). We made no directional predictions.

## RQ1: Publisher Information.

```{r total-articles}
# updated 
total_num <- 510334
health_num <- 445301
physical_num <- 228194
social_num <- 26652
life_num <- 307514
```

*Number of articles*. The total number of articles included in this analysis was `r total_num` including `r health_num` Health Sciences articles, `r physical_num`
Physical Sciences articles, `r social_num` Social Sciences articles, and
`r life_num` Life Sciences articles. Articles could be classified into
multiple categories. Figure \@ref(fig:fig-pub-time) shows the number of articles published across time for each of the four large subject areas.

```{r fig-pub-time, include = TRUE, fig.cap="Number of big-team science publications separated by four large subject areas across years."}
ggplot(DF$pub_time_count %>% 
         filter(SubjectArea != "Multidisciplinary") %>% 
         filter(Year < 2023), 
       aes(Year, count, color = SubjectArea)) + 
  theme_classic() + 
  geom_point() + 
  scale_color_brewer(palette = "Spectral", name = "Subject Area") + 
  ylab("Number of Articles Per Year") + 
  xlab("Publication Year") + 
  theme(legend.position = "bottom")
```

```{r total-journals}
# updated
journals <- 14924
health_journal <- 6559
physical_journal <- 5787
social_journal <- 2500
life_journal <- 4187
```

*Number of journals*. The number of distinct journals big team science articles were published
in was `r journals` with `r health_journal` journals in Health Sciences,
`r physical_journal` journals in Physical Sciences, `r social_journal`
journals in Social Sciences, and `r life_journal` journals in Life
Sciences. The descriptive statistics for the Source Normalized Impact
per Paper is presented the supplemental materials with a comparison for all papers.

## RQ2: Publication Information.

Publication interest area was summarized by the four large subject areas creating a word cloud plot of the total number of publications within the ASJCs. Figure \@ref(fig:fig-clouds) displays that the health sciences tends to publish within medicine and oncology, with a corresponding focus of cancer research and genetics for the life sciences. The physical sciences is mostly dominated by physics research, chemistry, and ecology. The BTS publications in the social sciences are mostly within psychology, education, and health.

```{r make-clouds, eval = FALSE, warning = FALSE, echo = FALSE, message = FALSE}
### NOTE YOU WILL NEED TO RUN THIS ###
#DF$pub_time_subject$ASJC <- factor(DF$pub_time_subject$ASJC)
pub_total <- DF$pub_time_subject %>% 
  group_by(SubjectArea, ASJC) %>% 
  summarize(Count_Size = sum(count), .groups = "keep") %>% 
  filter(SubjectArea != "Multidisciplinary") %>% 
  left_join(DF$ASJC_codes %>% select(Name, Code), by = c("ASJC" = "Code"))

# health
health_cloud <- pub_total %>% 
  filter(SubjectArea == "Health Sciences") %>% 
  ungroup() %>%
  arrange(desc(Count_Size)) %>% 
  ggplot() + 
  geom_text_wordcloud_area(aes(label = Name, size = Count_Size), area_corr_power = TRUE, rm_outside = TRUE, shape = "square", perc_step = .001, grid_margin = .01) + 
  theme_minimal() + 
  scale_size_area(max_size = 15)

ggsave("figure/health_cloud.png", dpi = 300, width = 8,
       height = 4, units = "in")

# physical
physical_cloud <- pub_total %>% 
  filter(SubjectArea == "Physical Sciences") %>% 
  ungroup() %>%
  arrange(desc(Count_Size)) %>% 
  ggplot() + 
  geom_text_wordcloud_area(aes(label = Name, size = Count_Size), area_corr_power = TRUE, rm_outside = TRUE, shape = "square", perc_step = .001, grid_margin = .01) + 
  theme_minimal() + 
  scale_size_area(max_size = 15)

ggsave("figure/physical_cloud.png", dpi = 300, width = 8,
       height = 4, units = "in")

# life
life_cloud <- pub_total %>% 
  filter(SubjectArea == "Life Sciences") %>% 
  ungroup() %>%
  arrange(desc(Count_Size)) %>% 
  ggplot() + 
  geom_text_wordcloud_area(aes(label = Name, size = Count_Size), area_corr_power = TRUE, rm_outside = TRUE, shape = "square", perc_step = .001, grid_margin = .01) + 
  theme_minimal() + 
  scale_size_area(max_size = 15)

ggsave("figure/life_cloud.png", dpi = 300, width = 8,
       height = 4, units = "in")

# social
social_cloud <- pub_total %>% 
  filter(SubjectArea == "Social Sciences") %>% 
  ungroup() %>%
  arrange(desc(Count_Size)) %>% 
  ggplot() + 
  geom_text_wordcloud_area(aes(label = Name, size = Count_Size), area_corr_power = TRUE, rm_outside = TRUE, shape = "square", perc_step = .001, grid_margin = .01) + 
  theme_minimal() + 
  scale_size_area(max_size = 15)

ggsave("figure/social_cloud.png", dpi = 300, width = 8,
       height = 4, units = "in")
```

```{r fig-clouds, include = TRUE, fig.cap = "Journal Areas for Big-Team Science Publications by Subject Area"}

health_cloud <- fig("figure/health_cloud.png")
physical_cloud <- fig("figure/physical_cloud.png") 
life_cloud <- fig("figure/life_cloud.png") 
social_cloud <- fig("figure/social_cloud.png")

health_cloud + ggtitle("Health Sciences") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
physical_cloud + ggtitle("Physical Sciences") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
life_cloud + ggtitle("Life Sciences") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
social_cloud + ggtitle("Social Sciences") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  plot_layout(ncol = 2, 
              nrow = 2)
```

## RQ3: Authors.

```{r total-authors}
# updated
total_authors <- 510334
avg_authors <- apa_num(49.30516877182394)
sd_authors <- apa_num(212.9786134634032)
min_authors <- 10
med_authors <- 18
max_authors <- 5568
```

The total number of unique authors across all publications was `r total_authors`. The mean number of authors per publication was *M* = `r avg_authors` (*SD* = `r sd_authors`, *Med* = `r med_authors`) with a range of `r min_authors` to `r max_authors`. The median and average number of authors by subject area are displayed in Figure \@ref(fig:fig-author-year). In general, the average and median number of authors increased over time, with the exception of the skew in the physical sciences. Interestingly, the effect in the physical sciences appears to be declining toward the general trends seen in other areas in the last few decades.

```{r fig-author-year, include = TRUE, fig.cap="Number of authors included on big-team science papers per year by subject area. Given the large skew in the data, the left panel presents the median number of authors per manuscript, and the right panel presents the average number of authors per manuscript by year."}
avg_author_time <- 
  DF$author_time %>% 
  filter(SubjectArea != "Multidisciplinary") %>% 
  # filter(Number_Papers >= 3) %>% 
  mutate(low = (Average_Num_Authors - as.numeric(SD_Num_Authors)), 
         high = (Average_Num_Authors + as.numeric(SD_Num_Authors)))

# ggplot(avg_author_time, aes(Year, Med_Num_Authors)) + 
#   theme_classic() + 
#   geom_point(aes(size = Number_Papers)) + 
#   geom_point(aes(Year, Average_Num_Authors, color = "red")) + 
#   #geom_pointrange(aes(ymin = low, ymax = high)) + 
#   facet_wrap(~SubjectArea) + 
#   ylab("Average Number of Authors") + 
#   xlab("Publication Year") + 
#   theme(legend.position = "bottom") + 
#   guides(color = "none") + 
#   scale_size_continuous(name = "Number Papers")

median_authors <- 
  ggplot(avg_author_time, aes(Year, Med_Num_Authors, color = SubjectArea)) + 
  theme_classic() + 
  geom_point() + 
    ylab("Median Number of Authors") + 
    xlab("Publication Year")

mean_authors <- 
  ggplot(avg_author_time, aes(Year, Average_Num_Authors, color = SubjectArea)) + 
  theme_classic() + 
  geom_point() + 
    ylab("Average Number of Authors") + 
    xlab("Publication Year")

median_authors + 
  scale_color_brewer(palette = "Spectral", name = "Subject Area") + 
  mean_authors + 
  scale_color_brewer(palette = "Spectral", name = "Subject Area") + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')
```

***Career Length***.

Figure \@ref(fig:fig-career) portrays the average career length for authors involved in BTS publications across years. Career length was defined as the year of first publication minus the current year, and higher numbers mean longer careers. To analyze trends over time, we  calculated the average career length for each publication (i.e., average author career lengths to create one score for each paper) and analyzed a regression analysis using career length to predict year of publication. In order to show variance between individuals, we calculated the standard deviation of career length for each publication and used variance as an additional predictor.

Negative career length slopes would indicate more young scholars in later years (i.e., lower average career length as time increases). Positive career length slopes would indicate older scholars in later years (i.e., higher average career length as time increases). Negative career variance slopes imply that variability decreases over the years, so the average career length is more homogeneous. Positive career length slopes imply that variability increases over the years, so the average career length is varied across individuals (i.e., different stages of scholars). Figure \@ref(fig:fig-heatmap) displays the results for all regression analyses to compare coefficient strength across and within hypothesis.

All values for this analyses were different from zero. The slopes for the average career length were negative for all four subject areas, indicating a trend toward younger scientist involvement over time for each area, with the strongest effect in the Physical sciences. The coefficient for variability in career length was also negative for each of the four subject areas with the highest in the Physical sciences and lowest in the Life Sciences. This result indicates a decrease in the variability of career lengths over time, likely from two sources: 1) more publications with more authors, thus, lowering variance estimations, and 2) more young scholars overall. The effect sizes for this analysis were surprisingly large ranging from $R^2$ to .25 to .47. All values and their confidence intervals can be found on our OSF page.

```{r fig-career, include = TRUE, fig.cap = "Average career length for big-team science authors. Larger dots indicate more variability in career length for authors by averaging the standard deviation in career length for each manuscript within a year. The data has been filtered to at least 10 publications in a year for this graph."}
ggplot(DF$career_time %>% 
         filter(SubjectArea != "Multidisciplinary") %>% 
         filter(num_papers >= 10), 
       aes(Year, avg_career, color = SubjectArea)) +
  theme_classic() + 
  geom_point(aes(size = sd_career)) +
  # geom_point() + 
  # facet_wrap(~SubjectArea) + 
  theme(legend.position = "bottom") + 
  ylab("Average Career Length") +
  xlab("Year of Publication") + 
  scale_color_brewer(palette = "Spectral", name = "Subject Area") +
  guides(size = "none")  
  #scale_size_continuous(name = "Number Papers")
```

```{r career-results, include = FALSE}

label <- c("Physical", "Social", "Health", "Life")
avg_b_values <- c(-0.692014, -0.365357, -0.451827, -0.508350)
avg_se_values <- c(0.002727, 0.005525, 0.001348, 0.001786)
sd_b_values <- c(-0.265538, -0.097108, -0.096730, -0.013958)
sd_se_values <- c(0.007018, 0.013409, 0.003451, 0.004523)
r2_values <- c(0.465446, 0.254312, 0.303781, 0.350923)
sample_size <- total_num

r2_low <- 1:length(r2_values)
r2_high <- 1:length(r2_values)
for (i in 1:length(r2_values)){
  temp <- ci.R2(R2 = r2_values[i], 1, sample_size)
  r2_low[i] <- temp$Lower.Conf.Limit.R2
  r2_high[i] <- temp$Upper.Conf.Limit.R2
}

graph_data <- data.frame(
  "label" = label,
  "estimate" = avg_b_values,
  "type" = "Coefficient",
  "low" = avg_b_values - qt(.025, sample_size, lower.tail = F)*avg_se_values,
  "high" = avg_b_values + qt(.025, sample_size, lower.tail = F)*avg_se_values) %>% 
    bind_rows(
    data.frame(
    "label" = label,
    "estimate" = sd_b_values,
    "type" = "SD Coefficient",
    "low" = sd_b_values - qt(.025, sample_size, lower.tail = F)*sd_se_values,
    "high" = sd_b_values + qt(.025, sample_size, lower.tail = F)*sd_se_values) 
    ) %>% 
  bind_rows(
    data.frame(  
      "label" = label, 
      "estimate" = r2_values,
      "type" = "Effect Size",
      "low" = r2_low,
      "high" = r2_high)
  )

graph_data$type <- factor(
  graph_data$type, 
  levels = c("Coefficient", "SD Coefficient", "Effect Size")
)

graph_data$analysis <- "Career"

# ggplot(graph_data, aes(label, estimate)) + 
#   geom_point() +
#   geom_pointrange(aes(ymin = low, ymax = high)) +
#   theme_classic() + 
#   facet_wrap(~type, labeller  = labeller(type = label_expressions(strip_labels))) + 
#   geom_hline(yintercept = 0) + 
#   coord_flip() + 
#   xlab("Subject Area") + 
#   ylab("")
```

```{r pub-results, include = FALSE}

label <- c("Physical", "Social", "Health", "Life")
avg_b_values <- c(0.004248, -0.027315, -0.021713, -0.009466)
avg_se_values <- c(0.000179, 0.001073, 0.000211, 0.000213)
sd_b_values <- c(-0.021552, 0.016810, 0.010139, -0.000378)
sd_se_values <- c(0.000241, 0.000921, 0.000198, 0.000235)
r2_values <- c(0.135965, 0.035968, 0.048999, 0.030748)
sample_size <- total_num

r2_low <- 1:length(r2_values)
r2_high <- 1:length(r2_values)
for (i in 1:length(r2_values)){
  temp <- ci.R2(R2 = r2_values[i], 1, sample_size)
  r2_low[i] <- temp$Lower.Conf.Limit.R2
  r2_high[i] <- temp$Upper.Conf.Limit.R2
}

graph_data_2 <- data.frame(
  "label" = label,
  "estimate" = avg_b_values,
  "type" = "Coefficient",
  "low" = avg_b_values - qt(.025, sample_size, lower.tail = F)*avg_se_values,
  "high" = avg_b_values + qt(.025, sample_size, lower.tail = F)*avg_se_values) %>% 
    bind_rows(
    data.frame(
    "label" = label,
    "estimate" = sd_b_values,
    "type" = "SD Coefficient",
    "low" = sd_b_values - qt(.025, sample_size, lower.tail = F)*sd_se_values,
    "high" = sd_b_values + qt(.025, sample_size, lower.tail = F)*sd_se_values) 
    ) %>% 
  bind_rows(
    data.frame(  
      "label" = label, 
      "estimate" = r2_values,
      "type" = "Effect Size",
      "low" = r2_low,
      "high" = r2_high)
  )

graph_data_2$type <- factor(
  graph_data_2$type, 
  levels = c("Coefficient", "SD Coefficient", "Effect Size")
)

graph_data_2$analysis <- "Publication"

# ggplot(graph_data, aes(label, estimate)) + 
#   geom_point() +
#   geom_pointrange(aes(ymin = low, ymax = high)) +
#   theme_classic() + 
#   facet_wrap(~type, labeller  = labeller(type = label_expressions(strip_labels))) + 
#   geom_hline(yintercept = 0) + 
#   coord_flip() + 
#   xlab("Subject Area") + 
#   ylab("")
```

```{r diversity-results, include = FALSE}

label <- c("Physical", "Social", "Health", "Life")
b_values <- c(0.002175, 0.033175, 0.016824, 0.008029)
se_values <- c(0.002578, 0.006039, 0.001150, 0.001476)
r2_values <- c(0.000006, 0.001542, 0.000708, 0.000170)
sample_size <- total_num

r2_low <- 1:length(r2_values)
r2_high <- 1:length(r2_values)
for (i in 1:length(r2_values)){
  temp <- ci.R2(R2 = r2_values[i], 1, sample_size)
  r2_low[i] <- temp$Lower.Conf.Limit.R2
  r2_high[i] <- temp$Upper.Conf.Limit.R2
}

graph_data_3 <- data.frame(
  "label" = label,
  "estimate" = b_values,
  "type" = "Coefficient",
  "low" = b_values - qt(.025, sample_size, lower.tail = F)*se_values,
  "high" = b_values + qt(.025, sample_size, lower.tail = F)*se_values) %>% 
  bind_rows(
    data.frame(  
      "label" = label, 
      "estimate" = r2_values,
      "type" = "Effect Size",
      "low" = r2_low,
      "high" = r2_high)
  )

graph_data_3$type <- factor(
  graph_data_3$type,
  levels = c("Coefficient", "Effect Size")
)

graph_data_3$analysis <- "Regions"

# ggplot(graph_data, aes(label, estimate)) + 
#   geom_point() +
#   geom_pointrange(aes(ymin = low, ymax = high)) +
#   theme_classic() + 
#   facet_wrap(~type, labeller  = labeller(type = label_expressions(strip_labels_2))) + 
#   geom_hline(yintercept = 0) + 
#   coord_flip() + 
#   xlab("Subject Area") + 
#   ylab("")
```

```{r fig-heatmap, include = TRUE, fig.cap="Heatmap results of regression analyses for career length, number of publications, and geopolitical diversity in region. The top figure represents all results together for comparison across analyses. The bottom row represents individual heatmaps for each hypothesis to distinguish small differences between subject areas for those research questions. Non-significant results are indicated with NS on the plot."}
graph_overall <- graph_data %>% 
  bind_rows(graph_data_2) %>% 
  bind_rows(graph_data_3) %>% 
  mutate(type = factor(type, levels = c("Effect Size", "SD Coefficient", "Coefficient"))) %>% 
  mutate(Estimate = estimate) %>% 
  mutate(heat_label = ifelse(
    low < .00001 & high > .00001, "NS", ""
  ))

heat_results <- ggplot(graph_overall, aes(label, type, fill = Estimate)) + 
  geom_tile(color = "white", lwd = 1.5, linetype = 1) +
  coord_fixed() + 
  scale_y_discrete(labels  = c(TeX("$R^2$"), TeX("$b_{SD}$"), TeX("$b$"))) + 
  ylab("") + 
  xlab("") + 
  facet_wrap(~analysis) + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 60, vjust = .5), 
        legend.position = "right", 
        legend.key.size = unit(.75, 'cm')) + 
  scale_fill_distiller(palette = "Spectral") + 
  geom_text(aes(label=heat_label))

# heat_results

heat_results_1 <- ggplot(graph_overall %>% 
                           filter(analysis == "Career"), 
                         aes(label, type, fill = Estimate)) + 
  geom_tile(color = "white", lwd = 1.5, linetype = 1) +
  coord_fixed() + 
  scale_y_discrete(labels  = c(TeX("$R^2$"), TeX("$b_{SD}$"), TeX("$b_M$"))) + 
  ylab("") + 
  xlab("") + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 60, vjust = .5), 
        legend.position = "right") + 
  scale_fill_distiller(palette = "RdBu") + 
  ggtitle("Career")  + 
  geom_text(aes(label=heat_label))

heat_results_2 <- ggplot(graph_overall %>% 
                           filter(analysis == "Publication"), 
                         aes(label, type, fill = Estimate)) + 
  geom_tile(color = "white", lwd = 1.5, linetype = 1) +
  coord_fixed() + 
  scale_y_discrete(labels  = c(TeX("$R^2$"), TeX("$b_{SD}$"), TeX("$b_M$"))) + 
  ylab("") + 
  xlab("") + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 60, vjust = .5), 
        legend.position = "right") + 
  scale_fill_distiller(palette = "PuOr") + 
  ggtitle("Publication") + 
  geom_text(aes(label=heat_label))

heat_results_3 <- ggplot(graph_overall %>% 
                           filter(analysis == "Regions"), 
                         aes(label, type, fill = Estimate)) + 
  geom_tile(color = "white", lwd = 1.5, linetype = 1) +
  coord_fixed() + 
  scale_y_discrete(labels  = c(TeX("$R^2$"), TeX("$b$"))) + 
  ylab("") + 
  xlab("") + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 60, vjust = .5), 
        legend.position = "right") + 
  scale_fill_distiller(palette = "PiYG") +
  ggtitle("Regions") + 
  geom_text(aes(label=heat_label))

layout <- "
AAA
BCD
"
suppressWarnings(heat_results + 
  heat_results_1 + heat_results_2 + heat_results_3 + 
  plot_layout(design = layout))
```

***Institution***.

```{r total-aff}
# updated
affs <- 463876
```

The total number of unique affiliation across all papers was `r affs`.

***Publication Metrics***.

-   We will report descriptive statistics on the total number of
    publications and h-index for individuals overall.
    -   Do this for each unique person and report averages
    -   Do this for each publication and create an average for each
        publication

```{r pub-hindex}
avg_pub_count_author	<- apa_num(38.368910766896065)
std_pub_count_author <- apa_num(102.53740454680472)

avg_pub_count_paper <- apa_num(162.50184385336053)
sd_pub_count_paper <- apa_num(155.1665533053519)
avg_SD_pub_count_paper <- apa_num(164.27249797983876)
SD_SD_pub_count_paper <- apa_num(127.20989934008757)

avg_HIndex <- apa_num(33.64589690529209)
std_HIndex <- apa_num(127.33565834550778)
med_HIndex <- apa_num(8.0)

avg_of_avg_HIndex <- apa_num(198.87318851400775)
std_of_avg_HIndex <- apa_num(248.7847084313958)
avg_of_std_HIndex <- apa_num(211.7985924228093)
std_of_std_HIndex <- apa_num(238.53141146314053)
med_of_med_HIndex <- apa_num(68.0)
```

The average number of publications by authors on big team sciences
papers is *M* = `r avg_pub_count_author` (*SD* =
`r std_pub_count_author`). The publication counts were averaged across
authors for each publication, and then these average publication counts
were averaged across publications *M* = `r avg_pub_count_paper` (*SD* =
`r sd_pub_count_paper`). The average variability (i.e., the average
standard deviation with authors of a manuscript) with publication counts
of a paper was $M_{SD}$ = `r avg_SD_pub_count_paper` ($SD_{SD}$ =
`r SD_SD_pub_count_paper`).

The same process was completed with *h*-index for each author and
publication. The average *h*-index for authors overall was *M* =
`r avg_HIndex` (*SD* = `r std_HIndex`, *Med* = `r med_HIndex`). The
average *h*-index for publications was *M* = `r avg_of_avg_HIndex` (*SD*
= `r std_of_avg_HIndex`), and the variability of *h*-index across
manuscripts was $M_{SD}$ = `r avg_of_std_HIndex` ($SD_{SD}$ =
`r std_of_std_HIndex`, $Med_{Med}$ = `r med_of_med_HIndex`).

-   Next, we will use the same analyses described in the career length
    section to analyze trends over time. An increasing slope over time
    indicates that individuals who are publishing more are more
    represented in BTS over time (i.e., increasing numbers of scholars
    with higher publication rates), while a negative slope indicates
    more researchers with less publications.
-   A positive slope for standard deviation indicates increasing
    variance over time (i.e., more diversity in the individual
    publication rates), while a negative slope would indicate less
    diversity in researchers over time. While publication rates do not
    represent value as a researcher, they are often used in hiring and
    promotion decisions, and we will use this variable as a proxy to
    gauge the diversity in scholars represented in big teams.
-   We will separate these by each of the four subject areas.

***Geopolitical Regions***.

-   We will present visualizations of the country information listed for
    authors, and we will discuss the areas of world in which authors
    generally come from, as well as the lowest representation of
    authors.

```{r geo-map}
# serbia has changed codes
# updated but needs work
DF$geo_country$affiliation_tag_country <- toupper(DF$geo_country$affiliation_tag_country)
DF$geo_country <- subset(DF$geo_country, affiliation_tag_country != "NONE")
DF$geo_country$affiliation_tag_country[DF$geo_country$affiliation_tag_country == "SCG"] <- "SRB"

DF$geo_country <- DF$geo_country %>% 
  group_by(affiliation_tag_country) %>% 
  summarize(count = sum(count),
            count_all = sum(all_count))

# convert country code to region code
# this will warn you about the ones that have multiple countries 
# create a world map 
world_map <- map_data(map = "world")
world_map$orig_region <- world_map$region
world_map$region <- iso.alpha(world_map$region, n = 3)
world_map <- subset(world_map, region != "ATA")

# maybe try binning
DF$geo_country$count_binned <- if_else(
  DF$geo_country$count >= 1000000, "1,000,000+", 
  if_else(
    DF$geo_country$count >= 500000 & DF$geo_country$count < 1000000, "500,000-999,999",
    if_else(
      DF$geo_country$count >= 100000 & DF$geo_country$count < 500000, "100,000-499,999",
      if_else(
        DF$geo_country$count >= 10000 & DF$geo_country$count < 100000, "10,000-99,999", 
        if_else(
          DF$geo_country$count < 10000 & DF$geo_country$count >= 5000, "5,000-9,999",
          if_else(
            DF$geo_country$count < 5000 & DF$geo_country$count >= 1000, "1,000-4,999",
            if_else(
              DF$geo_country$count < 1000 & DF$geo_country$count >= 100, "100-999",
              "< 100"
            )
          )
        )
      )
    )
  )
)

DF$geo_country$count_binned <- factor(DF$geo_country$count_binned, 
                                   levels = unique(DF$geo_country$count_binned))

DF$geo_country$count_binned <- factor(DF$geo_country$count_binned, 
                                      levels = c("< 100", 
                                                 "100-999", 
                                                 "1,000-4,999", 
                                                 "5,000-9,999", 
                                                 "10,000-99,999",
                                                 "100,000-499,999",
                                                 "500,000-999,999",
                                                 "1,000,000+"))

# map of binned data 
bin_country <- ggplot(DF$geo_country) +
  geom_map(aes(map_id = affiliation_tag_country, fill = count_binned), map = world_map) +
  geom_polygon(data = world_map, 
               aes(x = long, y = lat, group = group), 
               colour = 'black', fill = NA) + 
  theme_void() + 
  # scale_fill_manual(name = "Sample Size",
  #                   values = c( "black", "#323232", "#646464","#969696","#c8c8c8")) 
  scale_fill_brewer(name = "Sample Size", palette = "Spectral")

# Warning: Some values were not matched unambiguously: ANT, ATA, BUR, BUX, BYS, CAI, CSK, CZH, DDR, EUE, HVO, PCI, ROM, SUN, SUX, TWN, UNO, YMD, YUG, ZAR

# tree map
DF$geo_country$un_region_sub <- suppressWarnings(
  countrycode(
  sourcevar = DF$geo_country$affiliation_tag_country,
  origin = 'iso3c', 
  destination = 'un.regionsub.name')
)

DF$geo_country$un_region <- suppressWarnings(
  countrycode(
  sourcevar = DF$geo_country$affiliation_tag_country,
  origin = 'iso3c', 
  destination = 'un.region.name')
)

DF$geo_country$un_region[DF$geo_country$affiliation_tag_country == "TWN"] <- "Asia"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "TWN"] <- "Eastern Asia"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "ANT"] <- "Western Europe"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "BUR"] <- "Western Africa"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "BYS"] <- "Eastern Europe"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "CSK"] <- "Eastern Europe"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "DOR"] <- "Eastern Europe"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "HVO"] <- "Sub-Saharan Africa"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "PCI"] <- "Eastern Asia"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "ROM"] <- "Eastern Europe"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "SUN"] <- "Eastern Europe"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "YMD"] <- "Western Asia"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "YUG"] <- "Eastern Europe"
DF$geo_country$un_region_sub[DF$geo_country$affiliation_tag_country == "ZAR"] <- "Sub-Saharan Africa"

DF$geo_country <- DF$geo_country %>% 
  na.omit()

tree <- ggplot(DF$geo_country, aes(area = count, fill = count_binned,
               label = affiliation_tag_country, subgroup = un_region_sub)) +
  geom_treemap() +
  geom_treemap_subgroup_border(colour = "white", size = 5) +
  # geom_treemap_subgroup_text(place = "top", grow = TRUE,
  #                            alpha = 0.25, colour = "black",
  #                            fontface = "italic") +
  geom_treemap_text(colour = "black", place = "centre",
                    size = 15, grow = FALSE) +  
  # scale_fill_manual(name = "Sample Size",
  #                   values = c("#c8c8c8", "#969696", "#646464", "#323232", "black")) 
  scale_fill_brewer(name = "Sample Size", palette = "Spectral")
  # scale_fill_gradient(name = "Sample Size",
  #                     low = "#c8c8c8", 
  #                     high = "#323232") 
```

```{r geo-map-all}
# maybe try binning
DF$geo_country$count_binned_all <- if_else(
  DF$geo_country$count_all >= 1000000, "1,000,000+", 
  if_else(
    DF$geo_country$count_all >= 500000 & DF$geo_country$count_all < 1000000, "500,000-999,999",
    if_else(
      DF$geo_country$count_all >= 100000 & DF$geo_country$count_all < 500000, "100,000-499,999",
      if_else(
        DF$geo_country$count_all >= 10000 & DF$geo_country$count_all < 100000, "10,000-99,999", 
        if_else(
          DF$geo_country$count_all < 10000 & DF$geo_country$count_all >= 5000, "5,000-9,999",
          if_else(
            DF$geo_country$count_all < 5000 & DF$geo_country$count_all >= 1000, "1,000-4,999",
            if_else(
              DF$geo_country$count_all < 1000 & DF$geo_country$count_all >= 100, "100-999",
              "< 100"
            )
          )
        )
      )
    )
  )
)

DF$geo_country$count_binned_all <- factor(DF$geo_country$count_binned_all, 
                                   levels = unique(DF$geo_country$count_binned_all))

DF$geo_country$count_binned_all <- factor(DF$geo_country$count_binned_all, 
                                      levels = c("< 100", 
                                                 "100-999", 
                                                 "1,000-4,999", 
                                                 "5,000-9,999", 
                                                 "10,000-99,999",
                                                 "100,000-499,999",
                                                 "500,000-999,999",
                                                 "1,000,000+"))

# map of binned data 
bin_country_all <- ggplot(DF$geo_country) +
  geom_map(aes(map_id = affiliation_tag_country, fill = count_binned_all), map = world_map) +
  geom_polygon(data = world_map, 
               aes(x = long, y = lat, group = group), 
               colour = 'black', fill = NA) + 
  theme_void() + 
  # scale_fill_manual(name = "Sample Size",
  #                   values = c( "black", "#323232", "#646464","#969696","#c8c8c8")) 
  scale_fill_brewer(name = "Sample Size", palette = "Spectral")

tree_all <- ggplot(DF$geo_country, aes(area = count_all, fill = count_binned_all,
               label = affiliation_tag_country, subgroup = un_region_sub)) +
  geom_treemap() +
  geom_treemap_subgroup_border(colour = "white", size = 5) +
  # geom_treemap_subgroup_text(place = "top", grow = TRUE,
  #                            alpha = 0.25, colour = "black",
  #                            fontface = "italic") +
  geom_treemap_text(colour = "black", place = "centre",
                    size = 15, grow = FALSE) +  
  # scale_fill_manual(name = "Sample Size",
  #                   values = c("#c8c8c8", "#969696", "#646464", "#323232", "black")) 
  scale_fill_brewer(name = "Sample Size", palette = "Spectral")
  # scale_fill_gradient(name = "Sample Size",
  #                     low = "#c8c8c8", 
  #                     high = "#323232") 
```

```{r fig-map-both, include = TRUE, fig.cap="Geopolitical regions represented in big-team science publications versus all publications."}
bin_country + ggtitle("Big-Team Publication Map") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(legend.position = "none") + 
tree + ggtitle("Big-Team Publication Mosiac Chart") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  guides(fill = "none") + 
bin_country_all + ggtitle("All Publications Map") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  guides(fill = "none") + 
tree_all + ggtitle("All Publications Mosiac Chart") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  guides(fill = "none") + 
  plot_layout(ncol = 2, 
              nrow = 2) + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')

# need to do the legend 
```

-   To understand the change in representation diversity, we will
    summarize the total number of geopolitical regions for each paper.
    Using a linear model, we will examine if the number of regions
    present is predicted by the year of publication. Increasing
    diversity would be represented by a positive slope, while decreasing
    diversity would be represented by a negative slope.

-   Last, we will examine the differences in representation for
    corresponding author sets versus all other authors. For papers with
    10 to 49 authors, we will use the three first authors and the last
    author to compare against other authors. For 50 to 99 authors, five
    first authors plus last will be used, and for all papers with more
    than 100 authors, we will use ten first authors and the last author.
    We will calculate the frequencies of each of the UN Sub-Regions for
    first authors versus other authors, converting these values to
    proportions. Given the expected small sample sizes of these
    contingency tables, we will group together titles based on the year
    of publication (assuming at least 5 publications per year, these may
    be binned by 5-year or smaller increments to increase sample size).
    For each grouping, we will calculate the effect size of the
    differences in frequencies comparing first authors to all other
    authors. Since this data is categorical, we will use Cramer's *V* to
    represent the effect size. If the effect size includes zero in its
    confidence interval, this result will imply that first and all other
    authors represent the same pattern of UN Sub-Region diversity. Any
    confidence interval that does include zero represents a difference
    in diversity.

```{r small-effects, warning = FALSE}
DF$small_first_gp <- DF$small_first_gp %>% 
  filter(aff_un != "other") %>% 
  pivot_wider(id_cols = Year, 
              names_from = aff_un, 
              values_from = count,
              values_fill = 0)

DF$small_other_gp <- DF$small_other_gp %>% 
  filter(aff_un != "other") %>% 
  pivot_wider(id_cols = Year, 
              names_from = aff_un, 
              values_from = count,
              values_fill = 0) %>% 
  select(colnames(DF$small_first_gp))

small_effect <- data.frame(
  year = 1:length(unique(DF$small_first_gp$Year)),
  v = 1:length(unique(DF$small_first_gp$Year)),
  low = 1:length(unique(DF$small_first_gp$Year)),
  high = 1:length(unique(DF$small_first_gp$Year))
)
i <- 1
for (year in unique(DF$small_first_gp$Year)){
  temp_table <- DF$small_first_gp %>% filter(Year == year) %>% select(-Year) %>% 
    bind_rows(DF$small_other_gp %>% filter(Year == year) %>% select(-Year))

  col_total <- colSums(temp_table)
  temp_table <- temp_table %>% 
    select(names(col_total)[col_total > 0])
  temp <- suppressWarnings(chisq.test(as.matrix(temp_table)))

  tryCatch(mote_temp <- v.chi.sq(x2 = temp$statistic, n = sum(temp_table), 
                        r = nrow(temp_table), c = ncol(temp_table), 
                        a = .05), 
       warning = function(w) { 
         mote_temp <<- v.chi.sq(x2 = temp$statistic, n = sum(temp_table), 
                        r = nrow(temp_table), c = ncol(temp_table), 
                        a = .05)
         mote_temp$vlow <<- 0
  }
  )

  small_effect$year[i] <- year
  small_effect$v[i] <- mote_temp$v
  small_effect$low[i] <- mote_temp$vlow
  small_effect$high[i] <- mote_temp$vhigh
  small_effect$people[i] <- mote_temp$n
  i <- i + 1
}

small_effect$high[is.na(small_effect$high)] <- 0
small_effect <- 
  small_effect %>% 
  filter(v != Inf)

small_dot <- ggplot(small_effect, aes(year,v)) + 
  geom_point(aes(size = people)) + 
  theme_classic() + 
  geom_pointrange(aes(ymin = low, ymax = high)) + 
  xlab("Year") + 
  ylab(expression(phi)) + 
  scale_size_continuous(name = "Num Authors")

DF_decade <- DF$small_first_gp %>% 
  mutate(who = "F") %>% 
  bind_rows(
    DF$small_other_gp %>% 
      mutate(who = "O")
  ) %>% 
  mutate(Decade = ifelse(Year < 1990, "1980s", 
                         ifelse(Year < 2000, "1990s", 
                                ifelse(Year < 2010, "2000s", 
                                       ifelse(Year < 2020, "2010s", "2020s"))))) %>% 
  group_by(who, Decade) %>% 
  summarize_all(sum) %>% 
  pivot_longer(cols = `Western Europe`:Mirconesia,
               names_to = "Country", 
               values_to = "Count")

# North America, Africa (combine), NWS Europe, Eastern Asia, Latin America, Rest of Europe, Rest of Asia, Everything Else

DF_decade$Country_2 <- DF_decade$Country
DF_decade$Country_2 <- gsub(".*Africa.*", "Africa", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("Northern Europe|Southern Europe|Western Europe", "NWS Europe", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("South-Eastern Europe|Eastern Europe", "ESE Europe", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("Central Asia|South-Eastern Asia|Southern Asia|Western Asia", "Asia", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("Melanesia|Mirconesia|New zealand and Australia|Polynesia", "Other", DF_decade$Country_2)

DF_decade_sm <- DF_decade %>% 
  group_by(Country_2, Decade, who) %>% 
  summarize(Count = sum(Count), .groups = "keep")

small_author <- ggplot(DF_decade_sm, aes(who, Count, fill = Country_2)) + 
    geom_bar(position="fill", stat="identity", color = "black") + 
  theme_classic() + 
  facet_grid(~Decade) + 
  xlab("Author Type") + 
  ylab("Percent of Authors") + 
  scale_fill_brewer(name = "Region", palette = "Spectral")
```

```{r medium-effects, warning = FALSE}
DF$medium_first_gp <- DF$medium_first_gp %>% 
  filter(aff_un != "other") %>% 
  pivot_wider(id_cols = Year, 
              names_from = aff_un, 
              values_from = count,
              values_fill = 0)

DF$medium_other_gp <- DF$medium_other_gp %>% 
  filter(aff_un != "other") %>% 
  pivot_wider(id_cols = Year, 
              names_from = aff_un, 
              values_from = count,
              values_fill = 0) %>% 
  select(colnames(DF$medium_first_gp))

medium_effect <- data.frame(
  year = 1:length(unique(DF$medium_first_gp$Year)),
  v = 1:length(unique(DF$medium_first_gp$Year)),
  low = 1:length(unique(DF$medium_first_gp$Year)),
  high = 1:length(unique(DF$medium_first_gp$Year))
)
i <- 1
for (year in unique(DF$medium_first_gp$Year)){
  temp_table <- DF$medium_first_gp %>% filter(Year == year) %>% select(-Year) %>% 
    bind_rows(DF$medium_other_gp %>% filter(Year == year) %>% select(-Year))

  col_total <- colSums(temp_table)
  temp_table <- temp_table %>% 
    select(names(col_total)[col_total > 0])
  temp <- suppressWarnings(chisq.test(as.matrix(temp_table)))
  tryCatch(mote_temp <- v.chi.sq(x2 = temp$statistic, n = sum(temp_table), 
                        r = nrow(temp_table), c = ncol(temp_table), 
                        a = .05), 
       warning = function(w) { 
         mote_temp <<- v.chi.sq(x2 = temp$statistic, n = sum(temp_table), 
                        r = nrow(temp_table), c = ncol(temp_table), 
                        a = .05)
         mote_temp$vlow <<- 0
  }
  )
  medium_effect$year[i] <- year
  medium_effect$v[i] <- mote_temp$v
  medium_effect$low[i] <- mote_temp$vlow
  medium_effect$high[i] <- mote_temp$vhigh
  medium_effect$people[i] <- mote_temp$n
  i <- i + 1
}

medium_effect$high[is.na(medium_effect$high)] <- 0
medium_effect <- 
  medium_effect %>% 
  filter(v != Inf)

medium_dot <- ggplot(medium_effect, aes(year,v)) + 
  geom_point(aes(size = people)) + 
  theme_classic() + 
  geom_pointrange(aes(ymin = low, ymax = high)) + 
  xlab("Year") + 
  ylab(expression(phi)) + 
  scale_size_continuous(name = "Num Authors")

DF_decade <- DF$medium_first_gp %>% 
  mutate(who = "F") %>% 
  bind_rows(
    DF$medium_other_gp %>% 
      mutate(who = "O")
  ) %>% 
  mutate(Decade = ifelse(Year < 1990, "1980s", 
                         ifelse(Year < 2000, "1990s", 
                                ifelse(Year < 2010, "2000s", 
                                       ifelse(Year < 2020, "2010s", "2020s"))))) %>% 
  group_by(who, Decade) %>% 
  summarize_all(sum) %>% 
  pivot_longer(cols = `Western Europe`:Melanesia, 
               #### check this later ####
               names_to = "Country", 
               values_to = "Count")

# North America, Africa (combine), NWS Europe, Eastern Asia, Latin America, Rest of Europe, Rest of Asia, Everything Else

DF_decade$Country_2 <- DF_decade$Country
DF_decade$Country_2 <- gsub(".*Africa.*", "Africa", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("Northern Europe|Southern Europe|Western Europe", "NWS Europe", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("South-Eastern Europe|Eastern Europe", "ESE Europe", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("Central Asia|South-Eastern Asia|Southern Asia|Western Asia", "Asia", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("Melanesia|Mirconesia|New zealand and Australia|Polynesia", "Other", DF_decade$Country_2)

DF_decade_sm <- DF_decade %>% 
  group_by(Country_2, Decade, who) %>% 
  summarize(Count = sum(Count), .groups = "keep")

medium_author <- ggplot(DF_decade_sm, aes(who, Count, fill = Country_2)) + 
    geom_bar(position="fill", stat="identity", color = "black") + 
  theme_classic() + 
  facet_grid(~Decade) + 
  xlab("Author Type") + 
  ylab("Percent of Authors") + 
  scale_fill_brewer(name = "Region", palette = "Spectral")
```

```{r large-effects, warning = FALSE}
DF$large_first_gp <- DF$large_first_gp %>% 
  filter(aff_un != "other") %>% 
  pivot_wider(id_cols = Year, 
              names_from = aff_un, 
              values_from = count,
              values_fill = 0)

DF$large_other_gp <- DF$large_other_gp %>% 
  filter(aff_un != "other") %>% 
  pivot_wider(id_cols = Year, 
              names_from = aff_un, 
              values_from = count,
              values_fill = 0) %>% 
  select(colnames(DF$large_first_gp))

large_effect <- data.frame(
  year = 1:length(unique(DF$large_first_gp$Year)),
  v = 1:length(unique(DF$large_first_gp$Year)),
  low = 1:length(unique(DF$large_first_gp$Year)),
  high = 1:length(unique(DF$large_first_gp$Year))
)
i <- 1
for (year in unique(DF$large_first_gp$Year)){
  temp_table <- DF$large_first_gp %>% filter(Year == year) %>% select(-Year) %>% 
    bind_rows(DF$large_other_gp %>% filter(Year == year) %>% select(-Year))

  col_total <- colSums(temp_table)
  temp_table <- temp_table %>% 
    select(names(col_total)[col_total > 0])
  temp <- suppressWarnings(chisq.test(as.matrix(temp_table)))
  tryCatch(mote_temp <- v.chi.sq(x2 = temp$statistic, n = sum(temp_table), 
                        r = nrow(temp_table), c = ncol(temp_table), 
                        a = .05), 
       warning = function(w) { 
         mote_temp <<- v.chi.sq(x2 = temp$statistic, n = sum(temp_table), 
                        r = nrow(temp_table), c = ncol(temp_table), 
                        a = .05)
         mote_temp$vlow <<- 0
  }
  )
  large_effect$year[i] <- year
  large_effect$v[i] <- mote_temp$v
  large_effect$low[i] <- mote_temp$vlow
  large_effect$high[i] <- mote_temp$vhigh
  large_effect$people[i] <- mote_temp$n
  i <- i + 1
}

large_effect$high[is.na(large_effect$high)] <- 0
large_effect <- 
  large_effect %>% 
  filter(v != Inf)

large_dot <- ggplot(large_effect, aes(year,v)) + 
  geom_point(aes(size = people)) + 
  theme_classic() + 
  geom_pointrange(aes(ymin = low, ymax = high)) + 
  xlab("Year") + 
  ylab(expression(phi)) + 
  scale_size_continuous(name = "Num Authors")

ggplot(large_effect %>% 
         arrange(v) %>% 
         mutate(count = 1:nrow(large_effect)),
       aes(count, v)) + 
  geom_point(aes(size = people)) +
  geom_linerange(aes(ymin = low, ymax = high), alpha = .5) +
  theme_classic()  
  #ylab(TeX("Raw Effect: $t$ Log RT")) + 
  #xlab("Different Processing Scenarios") + 
  # geom_hline(yintercept = 0) + 
  #theme(text = element_text(size = 8))

DF_decade <- DF$large_first_gp %>% 
  mutate(who = "F") %>% 
  bind_rows(
    DF$large_other_gp %>% 
      mutate(who = "O")
  ) %>% 
  mutate(Decade = ifelse(Year < 1990, "1980s", 
                         ifelse(Year < 2000, "1990s", 
                                ifelse(Year < 2010, "2000s", 
                                       ifelse(Year < 2020, "2010s", "2020s"))))) %>% 
  group_by(who, Decade) %>% 
  summarize_all(sum) %>% 
  pivot_longer(cols = `Western Europe`:`Central Asia`, 
               #### check this later ####
               names_to = "Country", 
               values_to = "Count")

# North America, Africa (combine), NWS Europe, Eastern Asia, Latin America, Rest of Europe, Rest of Asia, Everything Else

DF_decade$Country_2 <- DF_decade$Country
DF_decade$Country_2 <- gsub(".*Africa.*", "Africa", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("Northern Europe|Southern Europe|Western Europe", "NWS Europe", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("South-Eastern Europe|Eastern Europe", "ESE Europe", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("Central Asia|South-Eastern Asia|Southern Asia|Western Asia", "Asia", DF_decade$Country_2)
DF_decade$Country_2 <- gsub("Melanesia|Mirconesia|New zealand and Australia|Polynesia", "Other", DF_decade$Country_2)

DF_decade_sm <- DF_decade %>% 
  group_by(Country_2, Decade, who) %>% 
  summarize(Count = sum(Count), .groups = "keep")

large_author <- ggplot(DF_decade_sm, aes(who, Count, fill = Country_2)) + 
    geom_bar(position="fill", stat="identity", color = "black") + 
  theme_classic() + 
  facet_grid(~Decade) + 
  xlab("Author Type") + 
  ylab("Percent of Authors") + 
  scale_fill_brewer(name = "Region", palette = "Spectral")
```

```{r author-gpe-figure, include = TRUE, fig.cap="A comparison of author affiliation geopolitical region across decades. F stands for first authors and O stands for other authors."}
small_author + ggtitle("10-49 Authors") + #theme(legend.position = "none") +
medium_author + ggtitle("50-99 Authors") + #theme(legend.position = "none") + 
large_author + ggtitle("100 + Authors") + #theme(legend.position = "none") + 
guide_area() + 
plot_layout(
  ncol = 2,
  nrow = 2, 
  guides = "collect")

ggsave("figure/author_country.png", width = 10, height = 7, 
       units = "in")

# fix the legend
```

```{r effect-gpe-figure, include = TRUE, fig.cap = "Effect size of the differences in representation for UN Regions for author affiliations in big-team science papers by year."}
small_dot + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(0,1), xlim = c(1980, 2025)) +
  ggtitle("10-49 Authors") + 
medium_dot + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(0,1), xlim = c(1980, 2025)) +
  ggtitle("50-99 Authors") +
large_dot + 
  theme(legend.position = "none") + 
  coord_cartesian(ylim = c(0,1), xlim = c(1980, 2025)) +
  ggtitle("100 + Authors") + 
guide_area() + 
plot_layout(
  ncol = 2,
  nrow = 2, 
  guides = "collect")

ggsave("figure/effect_country.png", width = 10, height = 7, 
       units = "in")
```

\newpage

# References

```{=tex}
\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
```

::: {#refs custom-style="Bibliography"}
:::

```{=tex}
\endgroup
```

\newpage

# (APPENDIX) Appendix {-}

```{r child = "appendix.Rmd"}
```